A Recurrence

### 2014-12-08 draft.tex
finished period 5 and 6

We can also perform this analysis for period $5$  
\begin{proposition}
If $F$ is of the form in equation ~\ref{eq:firstcase} and has period 5 and $f$ is a polynomial then $f(y) = cy+c^2$. 
\end{proposition}
\begin{proof}
First we must show that $a_5=a_0$ and $a_6=a_1$. Well 
\begin{equation*}
a_2 = \frac{cy+c^2}{x},
\end{equation*}
and so 
\begin{equation*}
a_3 = \frac{ c \frac{cy+c^2}{x} +c^2}{y} = \frac{c^2y + c^3+c^3x}{xy},
\end{equation*}
so
\begin{equation*}
a_4  = \frac{c \frac{cy + c^2+c^2x}{xy} +c^2}{\frac{cy+c^2}{x}} = \frac{c^2y+c^2+c^2x+c^2xy}{y(cy+c^2)} = \frac{c(c+x)}{y},
\end{equation*}
giving us 
\begin{align*}
a_5 &=\left(c \frac{c(c+x)}{y} + c^2\right) \frac{xy}{c^2y+c^3+c^3x} = \frac{c^2y+c^3+c^3x}{y} \frac{xy}{c^2y+c^3+c^3x}= x.
\end{align*}
We can finally calculate $a_6$ and find 
\begin{align*}
a_6 = (cx + c^2) \frac{y}{cx+c^2} = y
\end{align*}
Therefore $f(y)=cy+c^2$ is periodic with period 5 as desired. Now we would like to show that all polynomial $f$ must be of this form. So if we consider an arbitrary $f$ we know that the first few terms of the sequence will be 
\begin{equation*}
x,y, \frac{f(y)}{x}, \frac{f(f(y)/x)}{y}, \frac{xf( f(f(y)/x)/y)}{f(y)}, x, y, \ldots
\end{equation*}
Now using the recurrence relation and setting $a_5=a$ and $a_6=y$, we obtain the relations
\begin{align*}
y &= \frac{ f(x)f(y)}{xf(f(f(y)/x)/y)} \\
x &= \frac{y f( xf(f(f(y)/x)/y)/f(y)}{f(f(y)/x)}.
\end{align*}
This gives us 
\begin{align*}
 \frac{f(x)}{y} &= \frac{ xf(f(f(y)/x)/y)}{f(y)} \\
xf\left(\frac{f(y)}{x}\right) &= y f\left( \frac{ xf(f(f(y)/x)/y)}{f(y)}\right).
\end{align*}
We note that the right hand side of the first equation can be plugged into the right hand side of the second equation to give us 
\begin{equation}
\label{period5relation}
xf\left(\frac{f(y)}{x}\right) = y f\left(\frac{f(x)}{y}\right).
\end{equation}
Now if $f$ is some polynomial it takes the form $f(z) = b_n z^n + \cdots + b_1 z + b_0$, with $b_n \neq 0$. Plugging this definition into Equation \ref{period5relation} gives us 
\begin{align*}
&x\left( b_n \left( \frac{f(y)}{x}\right)^n + \cdots + b_1 \left( \frac{f(y)}{x}\right) + b_0 \right) = \\
& y\left( b_n \left( \frac{f(x)}{y}\right)^n + \cdots + b_1 \left( \frac{f(x)}{y}\right) + b_0 \right). 
\end{align*}
Since $x$ and $y$ can take any values, we can set $x=1$ and let $y$ be arbitrary. The largest exponent of $y$ on the left hand side is $n^n$ while the largest exponent of $y$ on the right hand side is $1$. Since $y$ we can make it arbitrarily large so that the term with the largest exponent dominates, this gives us that $n^n=1$ so $n=1$ and $f(z) =b_1 z + b_0$. Plugging this back into Equation \ref{period5relation} gives us 
\begin{equation*}
x\left(b_1 \left( \frac{b_1 y + b_0}{x} \right) + b_0 \right) = y\left(b_1 \left( \frac{b_1 x + b_0}{y} \right) + b_0 \right),
\end{equation*}
which simplifies to 
\begin{equation*}
b_1^2 y + b_0 x = b_1^2 x + b_0 y.
\end{equation*}
Setting $x=0$ we must have $b_0=b_1^2$. Therefore in order for the recurrence relation to have period 5 we must have $f(z)=cy+c^2$, as desired. 
\end{proof}

We can also perform this analysis for period $6$  
\begin{proposition}
If $F$ is of the form in equation ~\ref{eq:firstcase} and has period 6 and $f$ is a polynomial then $f(y) = cy$. 
\end{proposition}
\begin{proof} 
It is simple to see that $f(y)=cy$ produces a recurrence of period 6. Now we would like to show that all polynomial $f$ must be of this form. If we consider some arbitrary $f$ we know that we must have 
\begin{align*}
a_6 =x = \frac{f(a_5)}{a_4} \\ 
a_7 = y = \frac{f(a_6)}{a_5},
\end{align*}
and so $a_5 = \frac{f(x)}{y}$. This also gives us that 
\begin{equation*}
a_4 x = f\left( \frac{f(x)}{y} \right).
\end{equation*}
Noting that 
\begin{equation*}
a_4 = \frac{ x f(f(f(y)/x)/y)}{f(y)}, 
\end{equation*}
we obtain 
\begin{equation}
\label{period6}
 f\left( \frac{f(x)}{y} \right) = \frac{ x^2 f(f(f(y)/x)/y)}{f(y)}.
\end{equation}
Since $x$ and $y$ can take any values, we can set $y=1$ and let $x$ be arbitrary. Furthermore since $f$ is some polynomial it takes the form $f(z) = b_n z^n + \cdots + b_1 z +b_0$. Plugging these two into Equation \ref{period6} gives us 
\begin{equation*}
f(b_n x^n + \cdots + b_1 x + b_0) = x^2 f\left( b_n \frac{f(1)^n}{x^n} + \cdots + b_1 \frac{f(1)}{x} + b_0 \right).
\end{equation*}
We note that the largest exponent of $x$ on the left hand side is $n^n$ while the largest exponent of $y$ on the right hand side is $x^2$ if $b_0 \neq 0$. Since we can make $x$ arbitrarily large so that the term with the largest exponent dominates, this gives us that we must have $n^n=2$ if $b \neq 0$. This equation has no integer solutions and so we must have $b_0 =0$. Now if $b_1 \neq 0$ the largest exponent on the right hand side is $1$ and $n^n=1$ is solved by $n=1$ and so $f(z) = b_1 z$ as desired. 
\end{proof}

>>>>

Similarly, for period 5, we make the assumption that $f(y)$ is a polynomial in $y$.

\begin{proposition}
If $F(x,y) = f(y)/x$ and $f(y)$ is a polynomial in $y$, then the resulting sequence has period 5 if and only if $f(y) = cy + c^2$ for any constant $c$.
\end{proposition}

\begin{proof}
We omit the proof that this function indeed produces period 5. The proof simply involves writing out the sequence as given by the recurrence relation, and verifying that this works.

We show that if the sequnce has period 5, then $f(y)$ must be of this form. We know that the sequence must look like
\begin{equation*}
x,y, \frac{f(y)}{x}, \frac{f(f(y)/x)}{y}, \frac{xf( f(f(y)/x)/y)}{f(y)}, x, y, \dots .
\end{equation*}
We know that the sixth term must be $\frac{y f( xf(f(f(y)/x)/y)/f(y))}{f(f(y)/x)}$ and the seventh term must be $\frac{ f(x)f(y)}{xf(f(f(y)/x)/y)}$ due to the recurrence relation. This gives the equations
\begin{align*}
x &= \frac{y f( xf(f(f(y)/x)/y)/f(y))}{f(f(y)/x)}\\
y &= \frac{ f(x)f(y)}{xf(f(f(y)/x)/y)}.
\end{align*}
These equations can be rewritten as
\begin{align*}
xf\left(\frac{f(y)}{x}\right) &= y f\left( \frac{ xf(f(f(y)/x)/y)}{f(y)}\right) \\
\frac{f(x)}{y} &= \frac{ xf(f(f(y)/x)/y)}{f(y)}. 
\end{align*}
We note that the right hand side of the second equation can be plugged into the right hand side of the first equation to give us 
\begin{equation*}
xf\left(\frac{f(y)}{x}\right) = y f\left(\frac{f(x)}{y}\right).
\end{equation*}
Now if $f$ is some polynomial, it takes the form $f(z) = b_n z^n + \cdots + b_1 z + b_0$, with $b_n \neq 0$. Plugging this definition into the above equation gives us 
\begin{align*}
&x\left( b_n \left( \frac{f(y)}{x}\right)^n + \cdots + b_1 \left( \frac{f(y)}{x}\right) + b_0 \right) = \\
& y\left( b_n \left( \frac{f(x)}{y}\right)^n + \cdots + b_1 \left( \frac{f(x)}{y}\right) + b_0 \right). 
\end{align*}
Since $x$ and $y$ can take any values, we can set $x=1$ and let $y$ be arbitrary. The largest exponent of $y$ on the left hand side is $n^2$ while the largest exponent of $y$ on the right hand side is $1$. Since $y$ can be arbitrarily large, the term with the largest exponent dominates. For large $y$, the only way the two sides can be equal is if $n^2 = 1$, and so $n = 1$. This means that the polynomial $f(z)$ must be linear, so we write $f(z) =b_1 z + b_0$. Simplifying the above equations by assuming linearity gives
\begin{equation*}
x\left(b_1 \left( \frac{b_1 y + b_0}{x} \right) + b_0 \right) = y\left(b_1 \left( \frac{b_1 x + b_0}{y} \right) + b_0 \right),
\end{equation*}
which simplifies to 
\begin{equation*}
b_1^2 y + b_0 x = b_1^2 x + b_0 y.
\end{equation*}
Setting $x=0$ we find that $b_0=b_1^2$. Therefore in order for the recurrence relation to have period 5 we must have $f(z)=cy+c^2$, as desired. 
\end{proof}

\subsection{Period 6}

The last case we consider is period 6. Our approach for handling this case is similar to the one we use for period 5. We give the following result.

\begin{proposition}
If $F(x,y) = f(y)/x$ and $f(y)$ is a polynomial in $y$, then the resulting sequence has period 6 if and only if $f(y) = cy$ for any constant $c$.
\end{proposition}

\begin{proof}
It is not hard to see that $f(y) = cy$ produced a recurrence with period 6. The sequence is simply
\[ x, y, \frac{cy}{x}, \frac{c^2}{x}, \frac{c^2}{y}, \frac{cx}{y}, x,y,\dots .\]

Now we show that all polynomials $f(y)$ that produce period 6 must be of this form. Labeling the terms of the sequence as $a_0, a_1, a_2, \dots$, we get that 
\begin{align*}
a_6 =x = \frac{f(a_5)}{a_4} \\ 
a_7 = y = \frac{f(a_6)}{a_5},
\end{align*}
due to the periodicity condition. From this, we can deduce that $a_5 = \frac{f(x)}{y}$. This also gives us that 
\begin{equation*}
a_4 x = f\left( \frac{f(x)}{y} \right).
\end{equation*}
Noting that the recurrence relation gives
\begin{equation*}
a_4 = \frac{ x f(f(f(y)/x)/y)}{f(y)}, 
\end{equation*}
we obtain 
\begin{equation*}
 f\left( \frac{f(x)}{y} \right) = \frac{ x^2 f(f(f(y)/x)/y)}{f(y)}.
\end{equation*}
Since $x$ and $y$ can take any values, we can set $y=1$ and let $x$ be arbitrary. Furthermore since $f$ is some polynomial, it takes the form $f(z) = b_n z^n + \cdots + b_1 z +b_0$. Plugging these two into the above equation gives us 
\begin{equation*}
f(b_n x^n + \cdots + b_1 x + b_0) = x^2 f\left( b_n \frac{f(1)^n}{x^n} + \cdots + b_1 \frac{f(1)}{x} + b_0 \right).
\end{equation*}
We note that the largest exponent of $x$ in the polynomial on the left hand side is $n^2$ while the largest exponent of $y$ in the polynomial on the right hand side is $x^2$ if $b_0 \neq 0$. Since we can make $x$ arbitrarily large so that the term with the largest exponent dominates, this gives $n^2 = 2$ if $b \neq 0$. This equation has no integer solutions, so we must have $b_0 =0$. Now if $b_1 = 0$ (and $b_0 = 0$), then the highest exponent of $x$ on the right hand side is 0, and so $n^2 = 0$, implying that the entire polynomial is the zero polynomial. The function is not periodic in this case. This means the only case remaining is when  $b_1 \neq 0$ (and $b_0 = 0$). In this case, the same argument as before tells us that $n^2 = 1$, so $n = 1$. So the only possible polynomial is $f(z) = b_1 z$ as desired. 
\end{proof}
